---
title: "Homework6"
author: "Kevin Gnaster, Zach Josten, Cade Balvanz"
date: "3/2/2022"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
data(happy, package="productplots")
HAPPY <- readRDS("HAPPY.rds")
```

# Part 1

### Question 1

Kevin Gnaster

```{r}
HAPPY = replace(HAPPY, HAPPY == "IAP", NA)
HAPPY = replace(HAPPY, HAPPY == "DK", NA)
HAPPY = replace(HAPPY, HAPPY == "NA", NA)
```

### Question 2

Kevin Gnaster

```{r}
HAPPY = HAPPY %>% mutate(
  age = replace(AGE, AGE == "89 AND OLDER", 89),
  age = as.numeric(age)) %>% select(-AGE)

HAPPY = HAPPY %>% mutate(HAPPY = factor(HAPPY), 
                         SEX = factor(SEX),
                         MARITAL = factor(MARITAL),
                         DEGREE = factor(DEGREE),
                         FINRELA = factor(FINRELA),
                         HEALTH = factor(HEALTH),
                         PARTYID = factor(PARTYID),
                         POLVIEWS = factor(POLVIEWS))
str(HAPPY)
```

### Question 3

Kevin Gnaster

```{r}
HAPPY$DEGREE <- factor(HAPPY$DEGREE, levels = c("lt high school", "high school", "junior college", "bachelor", "graduate school"))

HAPPY$HEALTH <- factor(HAPPY$HEALTH, levels = c("POOR","FAIR","GOOD","EXCELLENT"))

HAPPY$FINRELA <- factor(HAPPY$FINRELA, levels = c("FAR BELOW AVERAGE","BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE","FAR ABOVE AVERAGE"))

HAPPY$POLVIEWS <- factor(HAPPY$POLVIEWS, levels = c("EXTREMELY LIBERAL","LIBERAL","SLIGHTLY LIBERAL","MODERATE","SLGHTLY CONSERVATIVE","CONSERVATIVE","EXTRMLY CONSERVATIVE"))
```
# Part 2

## Kevin Gnaster - Happiness vs Political Views / Sex exploration 

```{r}
ggplot(data = HAPPY, aes(x = POLVIEWS)) +
  geom_bar(aes(fill = HAPPY), position = "dodge")

ggplot(data = HAPPY, aes(x = SEX)) +
  geom_bar(aes(fill = HAPPY), position = "dodge")
```

For the first plot, we can see a pretty even distribution of happiness ratings to political views. All of them seem pretty close when taking into account proportions.

For the second plot, we can see that women seem to be a bit more unhappy than males.


## Zach Josten - Happiness vs Year / Finrela
```{r}
Year.2000 = HAPPY %>% dplyr::filter(as.integer(YEAR) >= 2000)
ggplot(Year.2000, aes(x = YEAR, fill = Year.2000$HAPPY)) + geom_bar()

ggplot(HAPPY, aes(x = FINRELA)) + geom_bar(aes(fill = HAPPY), position = "dodge") + coord_flip()


## Cade Balvanz - Happiness vs Degree/Health

```{r}
ggplot(data=HAPPY, aes(x=HEALTH)) +geom_bar(aes(fill=HAPPY),position = position_dodge())
```
As we can see in the first graph for the most part people are happy no matter what there health is, with lots of higher values of HAppy and Very Happy,  but there is a difference in the poor value with a lot more no to happys in comparison with the other graphs. 
```{r}
ggplot(data=HAPPY, aes(x=DEGREE)) +geom_bar(aes(fill=HAPPY),position = position_dodge())
```
No matter the degrees the mini histograms are all very similar with lots of pretty happy's, and secondly very happy. and the Not to happys and NA's are fairly low. 
```

